---
title: "R Notebook"
output: html_notebook
---

QTW Case Study 2
Due 6/2/2020

Q7 - Read files for female runners using functions from 2.3
Create a dataframe 
Might need to use createDF() and extractVariables()
Q10 - compare age distribution of female runners across all 14 years (1999-2012)  of the races. 
Use stuff like - Quantile quantile plots, boxplots, density curves
  But can use other techniques also/instead, have well explained charts/graphs
How does the distribution change over the years? Gradual or sharp change? (changepoint analysis)
additional/extra - model out where the age trend is going, speculate on reasons why, factors, forecasts, ect
?room to look at more? What’s the example business scenario this report will be addressing? A - more of an EDA, tell the stakeholder what they need to know about the data. 

ex, if the average age/other attribute of female runners holding steady at x, marketing divisions could use that age to tailor outreach, ect
ex, forecasted rates of participation mean [some sort of structural changes to accomodate those changes in future events]

```{r, echo = FALSE}
library(ggplot2)
library(dplyr)
library(tswge)
library(changepoint)
library(rvest)
library(XML)
library(chron) #string to time conversion

```

```{r}
#pagination loop?
years = seq(1999, 2012, 1)
#u0 = paste('http://www.cballtimeresults.org/performances?utf8=%E2%9C%93&section=10M&year=', years[i],'&division=Overall+Women', sep = '')
```

```{r}
#non pagination approach
u99 = 'http://cherryblossom.org/results/1999/cb99f.html'
u00 = 'http://cherryblossom.org/results/2000/Cb003f.htm'
u01 = 'http://www.cherryblossom.org/results/2001/oof_f.html'
u02 = 'http://www.cherryblossom.org/results/2002/ooff.htm'
u03 = 'http://www.cherryblossom.org/results/2003/CB03-F.HTM'
u04 = 'http://www.cherryblossom.org/results/2004/women.htm'
u05 = 'http://www.cherryblossom.org/results/2005/CB05-F.htm'
u06 = 'http://www.cherryblossom.org/results/2006/women.htm'
u07 = 'http://www.cherryblossom.org/results/2007/women.htm'
u08 = 'http://www.cherryblossom.org/results/2008/women.htm'
u09 = 'http://www.cherryblossom.org/results/2009/09cucb-F.htm'
u10 = 'http://www.cherryblossom.org/results/2010/2010cucb10m-f.htm'
u11 = 'http://cherryblossom.org/results/2011/2011cucb10m-f.htm'
u12 = 'http://cherryblossom.org/results/2012/2012cucb10m-f.htm'
```

```{r}
#time conversion function for chron, input string, return time format as double
time_convert <- function(x){
  if(nchar(x) <= 5)
    y <- chron(times = paste('00:', x, sep = ''))
  else
    y = chron(times = x)
  
return(y)
}
#chron(times = paste('00:', x))
```


```{r}
#define function to input URL, row_range, and output dataframe
#target is the url for html thing
#row_range is number of lines to actually grab, use format (ex) 4:2000 for start and end points
getlines99 <- function(target, row_range){
  doc = htmlParse(target)
  preNode = getNodeSet(doc, '//pre')
  txt = xmlValue(preNode[[1]])
  lines = strsplit(txt, '\\n')[[1]]
  lines = lines[row_range]
  
  df <- data.frame(empty = character() ,place = character(), div_tot = character(), fname = character(), surname = character(), age = integer(), hometown = character(), hometown_2 = character(), time = character(), pace = character()) #create empty dataframe to populate with rows
  
for(i in 1:length(lines)){
  dummy_value = 0 #filling with NA breaks the two part last name detection
  row_temp = as.list(strsplit(lines[i], '\\s+')[[1]])
  if(length(row_temp) == 1)
  {
  row_temp= append(row_temp, c('this', 'row', 'was', 'empty', 'must', 'have', 'over', 'shot', 'target'))
  }else
    {
    if(length(row_temp) == 7)
      {
      row_temp = append(row_temp, dummy_value, after = 2) #fill in for missing div/tot
      row_temp = append(row_temp, dummy_value, after = 5) #fill in for missing age
      row_temp = append(row_temp, dummy_value, after = 7) #fill in for missing state
      }
    if(length(row_temp) == 8) #if row is missing div, age, but NOT hometown2 (state)
      {
      row_temp = append(row_temp, dummy_value, after = 2) #fill in for missing div/tot
      row_temp = append(row_temp, dummy_value, after = 5) #fill in for missing age 
      }
    if(length(row_temp) == 9) #if row doesn't have a state/single name in hometown1
      {
      row_temp = append(row_temp, dummy_value, after = 7) #fill in for missing state
      }
    if(is.na(as.integer(row_temp[6])) == TRUE)
      { #if the 6th column is some 2 part last name, combine and drop
      row_temp[5] <- paste(row_temp[5], row_temp[6], sep = ' ')
      row_temp[7] <- NULL
      }
    if(length(row_temp) == 11) #if the hometown is two words
      {
      row_temp[7] <- paste(row_temp[7], row_temp[8], sep = ' ') #take two word hometown and combine into 1
      row_temp[8] <- NULL #remove extra item from hometown 1
      }
    if(length(row_temp) == 12) #if the hometown is THREE! words
      {
      row_temp[7] <- paste(row_temp[7], row_temp[8], row_temp[9], sep = ' ') #take two word hometown and combine into 1
      row_temp[8] <- NULL #remove extra item from hometown 1
      row_temp[9] <- NULL
    }
   }
  df_temp = t(as.data.frame(as.matrix(as.character(row_temp))))
  #print(df_temp) #diagnostic
  df <- rbind(df, df_temp)
}
  names(df) = c('empty', 'rank', 'division/total', 'first_name', 'surname', 'age', 'hometown1', 'hometown2','time', 'pace')
  return(df)
}
```


```{r, warning = FALSE}
#using function testing
df99 <- getlines99(u99, 4:2356) #obtain row_range by looking at lines input
```
```{r}
#write 1999 to csv
write.csv(df99, "")
```

```{r}
getlines00 <- function(target, row_range){
doc = htmlParse(target)
  preNode = getNodeSet(doc,'//font')
  txt = xmlValue(preNode[[4]])
  lines = strsplit(txt, '\r\n')[[1]]
  lines = lines[row_range]
  
 df <- data.frame(empty = character() ,place = character(), div_tot = character(), number = character(), fname = character(), surname = character(), age = integer(), hometown = character(), hometown_2 = character(), time = character(), net_time = character())
dummy_value = NA #filling with NA breaks the two part last name detection

for(i in 1:10){
   row_temp = as.list(strsplit(lines[i], '\\s+')[[1]])
  if(length(row_temp) == 1)
  {
  row_temp= append(row_temp, c('this', 'row', 'was', 'empty', 'must', 'have', 'over', 'shot', 'target'))
  }else
    {
    if(length(row_temp) == 10) #if row doesn't have a state/single name in hometown1
      {
      row_temp = append(row_temp, dummy_value, after = 8) #fill in for missing state
      }
    if(length(row_temp) == 12) #if the hometown is two words
      {
      row_temp[8] <- paste(row_temp[8], row_temp[9], sep = ' ') #take two word hometown and combine into 1
      row_temp[9] <- NULL #remove extra item from hometown 1
      }
   }
  df_temp = t(as.data.frame(as.matrix(as.character(row_temp))))
  #print(df_temp) #diagnostic
  df <- rbind(df, df_temp)
}

  names(df) = c('empty', 'rank', 'division/total', 'number', 'first_name', 'surname', 'age', 'hometown1', 'hometown2','time', 'net time')
  
  return(df)
}
```


```{r}
getlines00 <- function(target){
  
doc = htmlParse(target)
  preNode = getNodeSet(doc,'//font')
  txt = xmlValue(preNode[[4]])
  lines = strsplit(txt, '\r\n')[[1]]
  lines = lines[3:2168]
df <- data.frame(empty = character() ,place = character(), div_tot = character(), number = character(), fname = character(), surname = character(), age = integer(), hometown = character(), hometown_2 = character(), time = character(), net_time = character())
dummy_value = NA #filling with NA breaks the two part last name detection
for(i in 1:length(lines)){
   row_temp = as.list(strsplit(lines[i], '\\s+')[[1]])
  if(length(row_temp) == 1)
  {
  row_temp= append(row_temp, c('this', 'row', 'was', 'empty', 'must', 'have', 'over', 'shot', 'target'))
  }else
    {
    if(length(row_temp) == 10) #if row doesn't have a state/single name in hometown1
      {
      row_temp = append(row_temp, dummy_value, after = 8) #fill in for missing state
      }
    if(length(row_temp) == 12) #if the hometown is two words
      {
      row_temp[8] <- paste(row_temp[8], row_temp[9], sep = ' ') #take two word hometown and combine into 1
      row_temp[9] <- NULL #remove extra item from hometown 1
    }
    if(length(row_temp) == 13) #if the hometown is two words AND last name is two words
      {
      row_temp[6] <- paste(row_temp[6], row_temp[7], sep = ' ') #combine two word last name
      row_temp[7] <- NULL
      row_temp[8] <- paste(row_temp[8], row_temp[9], sep = ' ') #take two word hometown and combine into 1
      row_temp[9] <- NULL #remove extra item from hometown 1
    }
      if(length(row_temp) == 14) #if the hometown is THREE! words
      {
      row_temp[8] <- paste(row_temp[8], row_temp[9], row_temp[10], sep = ' ') #take two word hometown and combine into 1
      row_temp[9] <- NULL #remove extra item from hometown 1
      row_temp[10] <- NULL
    }
      
   }
  df_temp = t(as.data.frame(as.matrix(as.character(row_temp))))
  #print(df_temp) #diagnostic
  df <- rbind(df, df_temp)
}
  names(df) = c('empty', 'rank', 'division/total', 'number', 'first_name', 'surname', 'age', 'hometown1', 'hometown2','time', 'net time')
  return(df)
}
```


```{r}
df00 <- getlines00(u00)
```

```{r}
getlines01 <- function(target){
  
doc = htmlParse(target)
#doc = htmlParse(u01)
  preNode = getNodeSet(doc,'//pre')
  txt = xmlValue(preNode[[1]])
  lines = strsplit(txt, '\r\n')[[1]]
  lines = lines[4:2975]
df <- data.frame(empty = character(), place = character() ,number = character(), fname = character(), surname = character(), age = integer(), hometown = character(), hometown_2 = character(), time = character(), net_time = character())
dummy_value = NA 
cols = 10 #number of counted columns in original txt file, ex, place, number, fname, lname, age, ht1, ht2, time, time

for(i in 1:length(lines)){
#for(i in 2225:length(lines)){
   row_temp = as.list(strsplit(lines[i], '\\s+')[[1]])
  if(length(row_temp) == 1)
  {
  row_temp= append(row_temp, c(rep('empty row', cols)))
  }else
    {
    if(length(row_temp) == cols -1){ #if row is missing slot h2
      row_temp = append(row_temp, dummy_value, after = cols - 3)
    }
    if(length(row_temp) == cols -2){ #if row is missing both address 
      row_temp = append(row_temp, dummy_value, after = cols - 3)
      row_temp = append(row_temp, dummy_value, after = cols - 3)
    }
      if(length(row_temp) == cols+3 && is.na(as.integer(row_temp[(cols-2)])) == FALSE) #if the last name is THREEEEEEE! words #age will be in slot 8 
      {
      row_temp[cols-5] <- paste(row_temp[cols-5], row_temp[cols-4], row_temp[cols-3], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-3] <- NULL #remove extra item name
      row_temp[cols-4] <- NULL #remove extra item from name
    }
    if(length(row_temp) == (cols +1) && is.na(as.integer(row_temp[(cols-3)])) == FALSE) #if row has extra column AND its in the names area (age is in slot 7, detect by converting slot 7 to int and if it is NOT NA, then slot 7 contains age
      {
      row_temp[(cols-5)] = paste(row_temp[(cols-5)], row_temp[(cols-4)], sep = ' ') #combine 2nd and 3rd name
      row_temp[(cols-6)] <- NULL #drop 3rd name
    }
    else if(length(row_temp) == (cols +1) && is.na(as.integer(row_temp[(cols-4)])) == FALSE)#if row has extra column AND its in the hometown area, because age in slot 6
      {
      row_temp[(cols-3)] = paste(row_temp[(cols-3)], row_temp[(cols-2)], sep = ' ') #combine 2nd and 3rd address
      row_temp[(cols-2)] <- NULL #3rd address label
      } 
    if(length(row_temp) == cols+2) #if the hometown is two words AND last name is two words
      {
      row_temp[(cols-5)] <- paste(row_temp[(cols-5)], row_temp[(cols-6)], sep = ' ') #combine two word last name
      row_temp[(cols-6)] <- NULL
      row_temp[(cols-3)] <- paste(row_temp[(cols-3)], row_temp[(cols-2)], sep = ' ') #take two word hometown and combine into 1
      row_temp[(cols-2)] <- NULL #remove extra item from hometown 1
    }
      if(length(row_temp) == cols+3) #if the hometown is THREE! words
      {
      row_temp[cols-2] <- paste(row_temp[cols-2], row_temp[cols-1], row_temp[cols], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-1] <- NULL #remove extra item from hometown 1
      row_temp[cols] <- NULL
      }
  
      
   }
  df_temp = t(as.data.frame(as.matrix(as.character(row_temp))))
  #print(df_temp) #diagnostic
  df <- rbind(df, df_temp)
}
  names(df) = c('empty', 'rank', 'number', 'first_name', 'surname', 'age', 'hometown1', 'hometown2','time', 'net time')
  return(df)
}
```

```{r, warning = FALSE}
#warning false because of the large number of NA coercion checks
df01 <- getlines01(u01)
```

```{r}
#2002
getlines02 <- function(target){
doc = htmlParse(target)
#doc = htmlParse(u02)
  preNode = getNodeSet(doc,'//pre')
  txt = xmlValue(preNode[[1]])
  lines = strsplit(txt, '\r\n')[[1]]
  lines = lines[4:(length(lines)-1)]

c = 0
df = data.frame()
dummy_value = NA 
cols = 10 #number of counted columns in original txt file, ex, place, number, fname, lname, age, ht1, ht2, time, time

for(i in 1:length(lines)){
#for(i in 2408:length(lines)){
  row_temp = as.list(strsplit(lines[i], '\\s+')[[1]])
  if(length(row_temp) == 1)
  {
  row_temp= append(row_temp, c(rep('empty row', cols)))
  }else
    {
    if(length(row_temp) == cols -1){ #if row is missing slot h2
      row_temp = append(row_temp, dummy_value, after = cols - 3)
    }
    if(length(row_temp) == cols -2){ #if row is missing both address 
      row_temp = append(row_temp, dummy_value, after = cols - 3)
      row_temp = append(row_temp, dummy_value, after = cols - 3)
    }
      if(length(row_temp) == cols+2 && is.na(as.integer(row_temp[(cols-2)])) == FALSE) #if the last name is THREE! words #age will be in slot 8 
      {
      row_temp[cols-5] <- paste(row_temp[cols-5], row_temp[cols-4], row_temp[cols-3], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-3] <- NULL #remove extra item name
      row_temp[cols-4] <- NULL #remove extra item from name
      }
      if(length(row_temp) == cols+3 && is.na(as.integer(row_temp[(cols-2)])) == FALSE) #if the last name is 3 words AND city is two words
      {
      row_temp[cols-5] <- paste(row_temp[cols-5], row_temp[cols-4], row_temp[cols-3], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-3] <- NULL #remove extra item name
      row_temp[cols-4] <- NULL #remove extra item from name
      row_temp[cols-3] <- paste(row_temp[cols-3], row_temp[cols-2], sep = ' ')
      row_temp[cols-2] <- NULL
      }
      if(length(row_temp) == cols+3 && is.na(as.integer(row_temp[(cols-1)])) == FALSE) # IF THE LAST NAME IS FOUR WORDS LONG #age slot 9 
      {
      row_temp[cols-5] <- paste(row_temp[cols-5], row_temp[cols-4], row_temp[cols-3], row_temp[cols-2], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-2] <- NULL
      row_temp[cols-3] <- NULL #remove extra item name
      row_temp[cols-4] <- NULL #remove extra item from name
    }
    if(length(row_temp) == (cols +1) && is.na(as.integer(row_temp[(cols-3)])) == FALSE) #if row has extra column AND its in the names area (age is in slot 7, detect by converting slot 7 to int and if it is NOT NA, then slot 7 contains age
      {
      row_temp[(cols-5)] = paste(row_temp[(cols-5)], row_temp[(cols-4)], sep = ' ') #combine 2nd and 3rd name
      row_temp[(cols-4)] <- NULL #drop 3rd name
    }
    else if(length(row_temp) == (cols +1) && is.na(as.integer(row_temp[(cols-4)])) == FALSE)#if row has extra column AND its in the hometown area, because age in slot 6
      {
      row_temp[(cols-3)] = paste(row_temp[(cols-3)], row_temp[(cols-2)], sep = ' ') #combine 2nd and 3rd address
      row_temp[(cols-2)] <- NULL #3rd address label
      } 
    if(length(row_temp) == cols+2) #if the hometown is two words AND last name is two words
      {
      row_temp[(cols-5)] <- paste(row_temp[(cols-5)], row_temp[(cols-6)], sep = ' ') #combine two word last name
      row_temp[(cols-6)] <- NULL
      row_temp[(cols-3)] <- paste(row_temp[(cols-3)], row_temp[(cols-2)], sep = ' ') #take two word hometown and combine into 1
      row_temp[(cols-2)] <- NULL #remove extra item from hometown 1
    }
      if(length(row_temp) == cols+3) #if the hometown is THREE! words
      {
      row_temp[cols-2] <- paste(row_temp[cols-2], row_temp[cols-1], row_temp[cols], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-1] <- NULL #remove extra item from hometown 1
      row_temp[cols] <- NULL
      }
   }
  df_temp = t(as.data.frame(as.matrix(as.character(row_temp))))
  #print(df_temp) #diagnostic
  df <- rbind(df, df_temp)
  #c = c+1
  #print(c)
}
  names(df) = c('empty', 'rank', 'number', 'first_name', 'surname', 'age', 'hometown1', 'hometown2','net time', 'time')
  return(df)
}
```

```{r, warning = FALSE}
df02 <- getlines02(u02) #hnnggg nothing is the same formatttttt
#works but is red?? 
```


```{r}
#for 2003
parseLines03 <- function(lines, colnumber){
#feed this lines from diagnostic part
#col number in case the number of columns changes
c = 0
df = data.frame()
dummy_value = NA 
cols = colnumber #number of counted columns in original txt file, ex, place, number, fname, lname, age, ht1, ht2, time, time

for(i in 1:length(lines)){
#for(i in 2408:length(lines)){
  row_temp = as.list(strsplit(lines[i], '\\s+')[[1]])
  if(length(row_temp) == 1)
  {
  row_temp= append(row_temp, c(rep('empty row', cols)))
  }else
    {
    if(length(row_temp) == cols -1){ #if row is missing slot h2
      row_temp = append(row_temp, dummy_value, after = cols - 3)
    }
    if(length(row_temp) == cols -2){ #if row is missing both address 
      row_temp = append(row_temp, dummy_value, after = cols - 3)
      row_temp = append(row_temp, dummy_value, after = cols - 3)
    }
      if(length(row_temp) == cols -3 && is.na(as.integer(row_temp[(cols-9)]))== FALSE){ #if row is missing both address AND the div/tot, slot 2 will have a number instead of div/tot
      row_temp = append(row_temp, dummy_value, after = cols -10)
      row_temp = append(row_temp, dummy_value, after = cols - 3)
      row_temp = append(row_temp, dummy_value, after = cols - 3)
    }
      if(length(row_temp) == cols+2 && is.na(as.integer(row_temp[(cols-2)])) == FALSE) #if the last name is THREE! words #age will be in slot 8 
      {
      row_temp[cols-5] <- paste(row_temp[cols-5], row_temp[cols-4], row_temp[cols-3], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-3] <- NULL #remove extra item name
      row_temp[cols-4] <- NULL #remove extra item from name
      }
      if(length(row_temp) == cols+3 && is.na(as.integer(row_temp[(cols-2)])) == FALSE) #if the last name is 3 words AND city is two words
      {
      row_temp[cols-5] <- paste(row_temp[cols-5], row_temp[cols-4], row_temp[cols-3], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-3] <- NULL #remove extra item name
      row_temp[cols-4] <- NULL #remove extra item from name
      row_temp[cols-3] <- paste(row_temp[cols-3], row_temp[cols-2], sep = ' ')
      row_temp[cols-2] <- NULL
      }
      if(length(row_temp) == cols+3 && is.na(as.integer(row_temp[(cols-1)])) == FALSE) # IF THE LAST NAME IS FOUR WORDS LONG #age slot 9 
      {
      row_temp[cols-5] <- paste(row_temp[cols-5], row_temp[cols-4], row_temp[cols-3], row_temp[cols-2], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-2] <- NULL
      row_temp[cols-3] <- NULL #remove extra item name
      row_temp[cols-4] <- NULL #remove extra item from name
    }
    if(length(row_temp) == (cols +1) && is.na(as.integer(row_temp[(cols-3)])) == FALSE) #if row has extra column AND its in the names area (age is in slot 7, detect by converting slot 7 to int and if it is NOT NA, then slot 7 contains age
      {
      row_temp[(cols-5)] = paste(row_temp[(cols-5)], row_temp[(cols-4)], sep = ' ') #combine 2nd and 3rd name
      row_temp[(cols-4)] <- NULL #drop 3rd name
    }
    else if(length(row_temp) == (cols +1) && is.na(as.integer(row_temp[(cols-4)])) == FALSE)#if row has extra column AND its in the hometown area, because age in slot 6
      {
      row_temp[(cols-3)] = paste(row_temp[(cols-3)], row_temp[(cols-2)], sep = ' ') #combine 2nd and 3rd address
      row_temp[(cols-2)] <- NULL #3rd address label
      } 
    if(length(row_temp) == cols+2) #if the hometown is two words AND last name is two words
      {
      row_temp[(cols-5)] <- paste(row_temp[(cols-5)], row_temp[(cols-6)], sep = ' ') #combine two word last name
      row_temp[(cols-6)] <- NULL
      row_temp[(cols-3)] <- paste(row_temp[(cols-3)], row_temp[(cols-2)], sep = ' ') #take two word hometown and combine into 1
      row_temp[(cols-2)] <- NULL #remove extra item from hometown 1
    }
      if(length(row_temp) == cols+3) #if the hometown is THREE! words
      {
      row_temp[cols-2] <- paste(row_temp[cols-2], row_temp[cols-1], row_temp[cols], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-1] <- NULL #remove extra item from hometown 1
      row_temp[cols] <- NULL
      }
   }
  df_temp = t(as.data.frame(as.matrix(as.character(row_temp))))
  #print(df_temp) #diagnostic
  df <- rbind(df, df_temp)
  c = c+1
  #print(c)
}
  return(df)
}
```

```{r, warning = FALSE}
#2003
doc = htmlParse(u03)
  preNode = getNodeSet(doc,'//pre')
  txt = xmlValue(preNode[[1]])
  lines = strsplit(txt, '\r\n')[[1]]
  lines = lines[4:(length(lines)-2)]
  #split up this diagnostic part since it's been pretty different for every year so far
  #now div/tot is backkkk
  
df03 <- parseLines03(lines, 11)
names(df03) = c('empty', 'div/tot','rank', 'number', 'first_name', 'surname', 'age', 'hometown1', 'hometown2','net time', 'time')
```

```{r, warning = FALSE}
#2004
doc = htmlParse(u04)
  preNode = getNodeSet(doc,'//pre')
  txt = xmlValue(preNode[[1]])
  lines = strsplit(txt, '\r\n')[[1]]
  lines = lines[9:(length(lines))]
  #split up this diagnostic part since it's been pretty different for every year so far
  #now div/tot is backkkk
 
df04 <- parseLines03(lines, 11)
names(df04) = c('empty', 'div/tot','rank', 'number', 'first_name', 'surname', 'age', 'hometown1', 'hometown2','net time', 'time')
```


```{r}
parseLines05.1 <- function(lines, colnumber){
  #for when row configuration is place, div/tot, name, name, age, h1, h2, net time, guntime, pace
  #ncols = 11
c = 0
df = data.frame()
dummy_value = NA 
cols = colnumber #number of counted columns in original txt file, ex, place, number, fname, lname, age, ht1, ht2, time, time

for(i in 1:length(lines)){
#for(i in 2454:length(lines)){
  row_temp = as.list(strsplit(lines[i], '\\s+')[[1]])
  #row_temp = as.list(strsplit(lines[2], '\\s+')[[1]])
  if(length(row_temp) == 1)
  {
  row_temp= append(row_temp, c(rep('empty row', (cols-1))))
  }else
    {
      if(row_temp[length(row_temp)] == '!'){
      row_temp[length(row_temp)] <- NULL #remove that ! column
    }
    if(length(row_temp) == cols -1){ #if row is missing slot h2
      row_temp = append(row_temp, dummy_value, after = cols - 4)
    }
    if(length(row_temp) == cols -2){ #if row is missing both address 
      row_temp = append(row_temp, dummy_value, after = cols - 3)
      row_temp = append(row_temp, dummy_value, after = cols - 3)
    }
    if(length(row_temp) == cols -3 && is.na(as.integer(row_temp[(cols-9)]))== FALSE){ #if row is missing both address AND the div/tot, slot 2 will have a number instead of div/tot
      row_temp = append(row_temp, dummy_value, after = cols -10)
      row_temp = append(row_temp, dummy_value, after = cols - 3)
      row_temp = append(row_temp, dummy_value, after = cols - 3)
    }
    if(length(row_temp) == cols -4 && is.na(as.integer(row_temp[(cols-9)]))== FALSE){ #if row is missing division, age, h1 AND h2
      #look at slot to be rank
      row_temp = append(row_temp, dummy_value, after = cols -10)
      row_temp = append(row_temp, dummy_value, after = cols - 6)
      row_temp = append(row_temp, dummy_value, after = cols - 6)
       row_temp = append(row_temp, dummy_value, after = cols - 6)
    }
      if(length(row_temp) == cols+2 && is.na(as.integer(row_temp[(cols-2)])) == FALSE) #if the last name is THREE! words #age will be in slot 8 
      {
      row_temp[cols-5] <- paste(row_temp[cols-5], row_temp[cols-4], row_temp[cols-3], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-3] <- NULL #remove extra item name
      row_temp[cols-4] <- NULL #remove extra item from name
      }
      if(length(row_temp) == cols+3 && is.na(as.integer(row_temp[(cols-2)])) == FALSE) #if the last name is 3 words AND city is two words
      {
      row_temp[cols-5] <- paste(row_temp[cols-5], row_temp[cols-4], row_temp[cols-3], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-3] <- NULL #remove extra item name
      row_temp[cols-4] <- NULL #remove extra item from name
      row_temp[cols-3] <- paste(row_temp[cols-3], row_temp[cols-2], sep = ' ')
      row_temp[cols-2] <- NULL
      }
      if(length(row_temp) == cols+3 && is.na(as.integer(row_temp[(cols-1)])) == FALSE) # IF THE LAST NAME IS FOUR WORDS LONG #age slot 9 
      {
      row_temp[cols-5] <- paste(row_temp[cols-5], row_temp[cols-4], row_temp[cols-3], row_temp[cols-2], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-2] <- NULL
      row_temp[cols-3] <- NULL #remove extra item name
      row_temp[cols-4] <- NULL #remove extra item from name
    }
    if(length(row_temp) == (cols +1) && is.na(as.integer(row_temp[(cols-4)])) == FALSE) #if row has extra column AND its in the names area (age is in slot 7
      {
      row_temp[(cols-6)] = paste(row_temp[(cols-6)], row_temp[(cols-5)], sep = ' ') #combine 2nd and 3rd name
      row_temp[(cols-5)] <- NULL #drop 3rd name
    }
    else if(length(row_temp) == (cols +1) && is.na(as.integer(row_temp[(cols-5)])) == FALSE)#if row has extra column AND its in the hometown area, because age in slot 6
      {
      row_temp[(cols-4)] = paste(row_temp[(cols-4)], row_temp[(cols-3)], sep = ' ') #combine 2nd and 3rd address
      row_temp[(cols-3)] <- NULL #3rd address label
      } 
    if(length(row_temp) == cols+2) #if the hometown is two words AND last name is two words
      {
      row_temp[(cols-5)] <- paste(row_temp[(cols-5)], row_temp[(cols-6)], sep = ' ') #combine two word last name
      row_temp[(cols-6)] <- NULL
      row_temp[(cols-4)] <- paste(row_temp[(cols-4)], row_temp[(cols-3)], sep = ' ') #take two word hometown and combine into 1
      row_temp[(cols-3)] <- NULL #remove extra item from hometown 1
    }
      if(length(row_temp) == cols+3) #if the hometown is THREE! words
      {
      row_temp[cols-2] <- paste(row_temp[cols-2], row_temp[cols-1], row_temp[cols], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-1] <- NULL #remove extra item from hometown 1
      row_temp[cols] <- NULL
      }
   }
  df_temp = t(as.data.frame(as.matrix(as.character(row_temp))))
  #print(df_temp) #diagnostic
  df <- rbind(df, df_temp)
  c = c+1
  #print(c)
}
return(df)
}
```


```{r, warning = FALSE}
#2005
doc = htmlParse(u05)
  preNode = getNodeSet(doc,'//pre')
  txt = xmlValue(preNode[[1]])
  lines = strsplit(txt, '\r\n')[[1]]
  lines = lines[9:(length(lines))]
  lines[6] <-"    6    6/3146 Turena M_Johnson_Lane  29 Usa                  55:36   55:37  5:34 " #4names and no h2
  lines[678] <- "  678   23/217  Runner IIvViiIvO   51 Washington DC      1:25:53 1:27:30  8:45 " #not writing an exception for 6 part last name
 
df05 <- parseLines05.1(lines, 11)
names(df05) = c('empty', 'rank', 'div/tot', 'first_name', 'surname', 'age', 'hometown1', 'hometown2','net time', 'time', 'pace')
```


```{r}
parseLines06 <- function(lines){
  #for when row configuration is place, div/tot,number , name, age, h1, h2, net time, guntime, pace
  
c = 0
df = data.frame()
dummy_value = NA 
cols = 12#colnumber #number of counted columns in original txt file, ex, place, number, fname, lname, age, ht1, ht2, time, time, no more state abbreviation

#for(i in 1:length(lines)){
for(i in 1:length(lines)){
  row_temp = as.list(strsplit(lines[i], '\\s+')[[1]])
  #row_temp = as.list(strsplit(lines[2], '\\s+')[[1]])
  if(length(row_temp) == 1)
  {
  row_temp= append(row_temp, c(rep('empty row', (cols-1))))
  }else
    {
    if(row_temp[length(row_temp)] == '!'){
    row_temp[length(row_temp)] <- NULL #remove that ! column
    }
    if(length(row_temp) == (cols -1) && is.na(as.integer(row_temp[(cols-5)])) == FALSE){ #if row is missing slot h2
      row_temp = append(row_temp, dummy_value, after = cols - 4)
    }
    if(length(row_temp) == cols -2){ #if row is missing both address 
      row_temp = append(row_temp, dummy_value, after = cols - 3)
      row_temp = append(row_temp, dummy_value, after = cols - 3)
    }
    if(length(row_temp) == (cols +1) && is.na(as.integer(row_temp[(cols-4)])) == FALSE) #if row has extra column AND its in the names area (age is in slot 8 if h1 and h2 are already there
      {
      row_temp[(cols-6)] = paste(row_temp[(cols-6)], row_temp[(cols-5)], sep = ' ') #combine 2nd and 3rd name
      row_temp[(cols-5)] <- NULL #drop 3rd name
    }
    else if(length(row_temp) == (cols +1) && is.na(as.integer(row_temp[(cols-5)])) == FALSE)#if row has extra column AND its in the hometown area, because age in slot 6
      {
      row_temp[(cols-4)] = paste(row_temp[(cols-4)], row_temp[(cols-3)], sep = ' ') #combine 2nd and 3rd address
      row_temp[(cols-3)] <- NULL #3rd address label
    } 
      else if(length(row_temp) == (cols +1) && is.na(as.integer(row_temp[(cols-3)])) == FALSE){ #if the last name has three parts AND h2 is missing (look for age in slot 9)
        row_temp[cols-6] <- paste(row_temp[cols-6], row_temp[cols-5], row_temp[cols-4], sep = '-')
        row_temp[(cols-5):(cols-4)] <- NULL
        row_temp = append(row_temp, dummy_value, after = cols - 4)
      }
    if(length(row_temp) == cols+2) #if the hometown is two words AND last name is two words
      {
      row_temp[(cols-5)] <- paste(row_temp[(cols-5)], row_temp[(cols-6)], sep = ' ') #combine two word last name
      row_temp[(cols-6)] <- NULL
      row_temp[(cols-4)] <- paste(row_temp[(cols-4)], row_temp[(cols-3)], sep = ' ') #take two word hometown and combine into 1
      row_temp[(cols-3)] <- NULL #remove extra item from hometown 1
    }
      if(length(row_temp) == cols+2 && is.na(as.integer(row_temp[(cols-3)])) == FALSE) #if the last name is THREE! words #age will be in slot 9 
      {
      row_temp[cols-6] <- paste(row_temp[cols-6], row_temp[cols-5], row_temp[cols-4], sep = ' ') #take 3 word surname and combine 
      row_temp[cols-5] <- NULL #remove extra item from name
      row_temp[cols-5] <- NULL #remove extra item name
      }
      if(length(row_temp) == cols+3) #if the hometown is THREE! words
      {
      row_temp[cols-2] <- paste(row_temp[cols-2], row_temp[cols-1], row_temp[cols], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-1] <- NULL #remove extra item from hometown 1
      row_temp[cols] <- NULL
      }
   }
  df_temp = t(as.data.frame(as.matrix(as.character(row_temp))))
  #print(df_temp) #diagnostic
  df <- rbind(df, df_temp)
  c = c+1
  print(c)
}
return(df)
}
```


```{r, warning = FALSE}
#2006
doc = htmlParse(u06)
  preNode = getNodeSet(doc,'//pre')
  txt = xmlValue(preNode[[1]])
  lines = strsplit(txt, '\r\n')[[1]]
  lines = lines[9:(length(lines)-2)]
#the state name is gone from hometown
lines[3719] <- " 3719    NA    9423 Robin Hershey   NA     Potomac    NA     1:45:06 1:50:47  11:05   " #case with no division, age, or h2
 
df06 <- parseLines06(lines)
names(df06) = c('empty', 'rank', 'div/tot', 'number', 'first_name', 'surname', 'age', 'hometown1', 'hometown2','net time', 'time', 'pace')
#check!
```


```{r}
parseLines07 <- function(lines, diagnostics){
  #for when row configuration is place, div/tot,number , name, age, h1, h2, net time, guntime, pace
  #diagnostics = 'yes' if you want to print out the rows as they are processed. used to find exactly where the parser stopped
c = 0
check_row = diagnostics
df = data.frame()
dummy_value = NA 
cols = 12#colnumber #number of counted columns in original txt file, ex, place, number, fname, lname, age, ht1, ht2, time, time, no more state abbreviation

#for(i in 1:length(lines)){
for(i in 1:length(lines)){
  row_temp = as.list(strsplit(lines[i], '\\s+')[[1]])
  #row_temp = as.list(strsplit(lines[1159], '\\s+')[[1]])
  if(length(row_temp) == 1)
  {
  row_temp= append(row_temp, c(rep('empty row', (cols-1))))
  }else
    {
    if(row_temp[length(row_temp)-1] == '!'){
    row_temp[length(row_temp)-1] <- NULL #remove that ! column
    }
    if(length(row_temp) == (cols -1) && is.na(as.integer(row_temp[(cols-5)])) == FALSE){ #if row is missing slot h2
      row_temp = append(row_temp, dummy_value, after = cols - 4)
    }
      else if(length(row_temp) == (cols -1) && is.na(as.integer(row_temp[(cols-5)])) == TRUE){
        #if short one col and age is not in col 7, add na for age
        row_temp = append(row_temp, dummy_value, after = (cols - 6))
      }
    if(length(row_temp) == cols -2){ #if row is missing both address 
      row_temp = append(row_temp, dummy_value, after = cols - 3)
      row_temp = append(row_temp, dummy_value, after = cols - 3)
    }
    if(length(row_temp) == (cols +1) && is.na(as.integer(row_temp[(cols-4)])) == FALSE) #if row has extra column AND its in the names area (age is in slot 8 if h1 and h2 are already there
      {
      row_temp[(cols-6)] = paste(row_temp[(cols-6)], row_temp[(cols-5)], sep = ' ') #combine 2nd and 3rd name
      row_temp[(cols-5)] <- NULL #drop 3rd name
    }
    else if(length(row_temp) == (cols +1) && is.na(as.integer(row_temp[(cols-5)])) == FALSE)#if row has extra column AND its in the hometown area, because age in slot 6
      {
      row_temp[(cols-4)] = paste(row_temp[(cols-4)], row_temp[(cols-3)], sep = ' ') #combine 2nd and 3rd address
      row_temp[(cols-3)] <- NULL #3rd address label
    } 
      else if(length(row_temp) == (cols +1) && is.na(as.integer(row_temp[(cols-3)])) == FALSE){ #if the last name has three parts AND h2 is missing (look for age in slot 9)
        row_temp[cols-6] <- paste(row_temp[cols-6], row_temp[cols-5], row_temp[cols-4], sep = '-')
        row_temp[(cols-5):(cols-4)] <- NULL
        row_temp = append(row_temp, dummy_value, after = cols - 4)
      }
    if(length(row_temp) == cols+2) #if the hometown is two words AND last name is two words
      {
      row_temp[(cols-5)] <- paste(row_temp[(cols-5)], row_temp[(cols-6)], sep = ' ') #combine two word last name
      row_temp[(cols-6)] <- NULL
      row_temp[(cols-4)] <- paste(row_temp[(cols-4)], row_temp[(cols-3)], sep = ' ') #take two word hometown and combine into 1
      row_temp[(cols-3)] <- NULL #remove extra item from hometown 1
    }
      if(length(row_temp) == cols+2 && is.na(as.integer(row_temp[(cols-3)])) == FALSE) #if the last name is THREE! words #age will be in slot 9 
      {
      row_temp[cols-6] <- paste(row_temp[cols-6], row_temp[cols-5], row_temp[cols-4], sep = ' ') #take 3 word surname and combine 
      row_temp[cols-5] <- NULL #remove extra item from name
      row_temp[cols-5] <- NULL #remove extra item name
      }
      if(length(row_temp) == cols+3) #if the hometown is THREE! words
      {
      row_temp[cols-2] <- paste(row_temp[cols-2], row_temp[cols-1], row_temp[cols], sep = ' ') #take two word hometown and combine into 1
      row_temp[cols-1] <- NULL #remove extra item from hometown 1
      row_temp[cols] <- NULL
      }
   }
  df_temp = t(as.data.frame(as.matrix(as.character(row_temp))))
  #print(df_temp) #diagnostic
  df <- rbind(df, df_temp)
  if(check_row == "yes"){
    c = c+1
    print(c)
  }
}
return(df)
}
```

```{r, warning = FALSE}
#2007
doc = htmlParse(u07)
  preNode = getNodeSet(doc,'//pre')
  txt = xmlValue(preNode[[1]])
  lines = strsplit(txt, '\r\n')[[1]]
  lines = lines[8:(length(lines)-2)]
###the state name is back... 
  #not time efficent to write new rules if the format is going to change so much
lines[346] <- "  346   NA      1973 Chris Mickeever    NA       Plymouth_Meetin PA 1:17:02   7:43     47:26 "
lines[489] <- "  489   138/1659    2518 Becky Carlson_St_Clair 28 Silver_Spring MD   1:19:29   7:57     49:49 "
lines[3345] <- " 3345   315/519    13141 Lilly Sehayek          40 NeW_Hyde_Park NY  1:39:43   9:59   1:03:16 "
df07 <- parseLines07(lines, diagnostics = 'no')
#df07 <- parseLines07(lines[3344:length(lines)])
names(df07) = c('empty', 'rank', 'div/tot', 'number', 'first_name', 'surname', 'age', 'hometown1', 'hometown2','net time', 'time', 'pace')
#seems to work
```


```{r}
#2008
#MORE COLUMNS WHY

doc = htmlParse(u08)
  preNode = getNodeSet(doc,'//pre')
  txt = xmlValue(preNode[[1]])
  lines = strsplit(txt, '\r\n')[[1]]
  lines = lines[9:(length(lines)-2)]
  
```

```{r}
#debugging
#c to see what row the parser stopped on
row_temp = as.list(strsplit(lines[(1161)], '\\s+')[[1]])
t(row_temp)

```

```{r}
if(row_temp[length(row_temp)] == '!'){
  #substr(lines[1], 1, (nchar(lines[1])-3)) #takes off the ! column
  row_temp[length(row_temp)] <- NULL
}

```

```{r}
doc = htmlParse(u02)
  preNode = getNodeSet(doc,'//pre')
  txt = xmlValue(preNode[[1]])
  lines = strsplit(txt, '\r\n')[[1]]
  lines = lines[4:length(lines)]
  row_temp = as.list(strsplit(lines[1], '\\s+')[[1]])
  t(row_temp)
  
```














