---
title: "case study 4 v3"
output: html_notebook
---


Lance Dacy, Reanan McDaniel, Shawn Jung, Jonathan Tan

```{r}
#analysis packages
library(tswge)
library(changepoint)
library(tseries)
library(DataExplorer)
library(nnfor)

#data manipulation
library(dplyr)
library(lubridate)
library(tibbletime)

#visualization
library(ggplot2)
library(cowplot)
library(kableExtra)

#data source
library(cdcfluview)
```

import data from... ??? https://apps.who.int/flumart/Default?ReportNo=12
query below is SE Asia from 1995 to 2019
```{r}
raw_data  <- read.csv('D:/SMU/DS 7333 Quantify The World/FluNetInteractiveReport(SE_Asia_1995-2019).csv')
```
Column guide for WHO data csv columns
SPEC_RECIEVED_NB - Specimens recieved by the WHO
SPEC_PROCESSED_NB - specimens processed by the WHO
AH1 - all A type influenza viruses categorized by subtype
A(H1N1pdm09
AH3
AH5
ANOTSUBTYPED
INF(A) - total of all influenza A viruses detected
BYAMAGATA - B(Yamagata lineage) - various types of influenze B type viruses
BVICTORIA - B(Victoria Lineage)
BNOTDETERMINED
INF_B - total influenze B viruses 
ALL_INF - number of influenze positive viruses
ALL_INF2 - number of influenze negative viruses
TITLE - ili activity? most have no report as the column

will have to look more into the details, but it looks like this dataset is recording the number of virus strains, not individual cases of that virus. specimens recieved/processed usually matches up with INF + or -, indicating that those specimens were all tested and sorted into either yes, it was an influenza virus, or no, it was not. 

We'll pick S asia, India for total influenza cases

```{r}
S_Asia <- raw_data[which(raw_data$FLUREGION == 'Southern Asia'), ] #select rows from SE asia
tlist <- S_Asia[complete.cases(S_Asia$ALL_INF), ] #select only rows that aren't missing data in the all_inf column
t_ind <- tlist[which(tlist['ï..Country'] == 'India'), ]
t_ind <- t_ind[which(t_ind$Year == 2017| t_ind$Year == 2018 | t_ind$Year == 2019), ] #select 3 years
ttime = t_ind$SDATE %>% as.Date("%m/%d/%Y") #store the timeline as date object vector
#it looks like all the weeks fall on the same SDATE's, so it should be safe to use the same date vector for different countries.

ggplot(t_ind) +
  geom_line(aes(x = ttime, y = ALL_INF, col = 'Total Influenza Cases'))+
  #geom_line(aes(x = ttime, y = AH1, col = 'AH1'))+
  #geom_line(aes(x = ttime, y = AH1N12009, col = '2009 Swine Flu'))+
  #geom_line(aes(x = ttime, y = AH3, col = 'AH3'))+
  #geom_line(aes(x = ttime, y = AH5, col = 'AH5'))+
  #geom_line(aes(x = ttime, y = ANOTSUBTYPED, col = 'Untyped Inf A'))+
  geom_line(aes(x = ttime, y = INF_A, col = 'total type A'))+
  #geom_line(aes(x = ttime, y = BYAMAGATA, col = 'Yamagata STRAIN B'))+
  #geom_line(aes(x = ttime, y = BVICTORIA, col = 'Victoria Strain B'))+
  #geom_line(aes(x = ttime, y = BNOTDETERMINED, col = 'Untyped Inf B'))+
  geom_line(aes(x = ttime, y = INF_B, col = 'total type B')) +
  scale_y_continuous()
```

diagnostics, stationarity
```{r}
tdata = t_ind$ALL_INF
#wrap objects in invisible() to hide the long console output and leave only the graph
invisible(plotts.sample.wge(tdata))
#stationairity
adf.test(tdata)
pacf(tdata)
#changepoint
par(mfrow = c(2,1))
v2 <- tdata
v2.pelt <- cpt.meanvar(v2, test.stat = 'Poisson', method = "PELT")
plot(v2.pelt, main = 'PELT changes in variance')
cpts.ts(v2.pelt)

v2.bs <- cpt.meanvar(v2, test.stat = 'Poisson', method = 'BinSeg')
plot(v2.bs, cpt.width = 3, main = 'BinSeg changes in Mean Variance')
cpts.ts(v2.bs)
```
augmented dicky fuller test yields -4.97, less than p-value of 0.01; reject null; insufficient evidency to claim time series is NOT stationary
basically it is stationary


feed into modeling block 
```{r}
#model fitting with tdata
pspan = 0:10 #range of values to look for possible p and q coefficients for AR and MA
qspan = 0:8
difference = 0
aic_results <- aic5.wge(tdata, p = pspan, q = qspan)
bic_results <- aic5.wge(tdata, p = pspan, q = qspan, type = 'bic')
aic_results
bic_results

m1 = est.arma.wge(tdata, p = aic_results[1, 1], q = aic_results[1, 2], factor = TRUE) #feed top AIC or BIC into arma estimation to generate coefficients
m1$phi
m1$theta

#simple single forecast and test
weeks_compare = 24 #how many weeks to reserve for testing : 6months 
f1 <- tdata[1:(length(tdata)-weeks_compare)] %>% fore.aruma.wge(phi = m1$phi, theta = m1$theta, d = 0, n.ahead = weeks_compare)
mse <- mean((tdata[((length(tdata)-weeks_compare)+1):(length(tdata))] - f1$f)^2)
paste('Mean Squared Error: ', mse)
#additional metrics
a_metrics <- accuracy(f1$f, tdata[((length(tdata)-weeks_compare)+1):(length(tdata))] )

#change x and y line names to match whatever dataframe you've stored the time series in
timeFrame <- data.frame(date = ttime, inf_cases = tdata)
g3 <- ggplot(timeFrame)+
  geom_line(aes(x = date, y = tdata, color = 'black'), size = 0.5)+
  geom_line(aes(x = date, y = c(rep(NA, length(tdata)-weeks_compare), f1$f), color = 'red'), size= 1) + 
  scale_color_discrete(name = "total_patients", labels = c('actual', 'predicted')) +
  ggtitle(paste('ARIMA (',aic_results[1, 1],',',aic_results[1, 2],',',difference,')', 'Forecast of ', weeks_compare, ' weeks'), subtitle = paste('Mean Squared Error = ', mse))
g3
```

```{r}
#updated with step size,how much the "window" moves per interval 
#rolling window ase: apply model predictions to small segments of the given data and aggregate metrics
trainingSize = 24
horizon = 4
step_size = 8
n_windows = round((length(tdata)-(trainingSize + horizon))/step_size) #number of whole windows of training size + horizon that will fit into the given data, divided by step size
ASEHolder = numeric()

fcastHolder = matrix(nrow = n_windows, ncol = horizon) #create matrix that has the number of rows and columns corresponding to number of windows and length of forecast interval
fUpperHolder = matrix(nrow = n_windows, ncol = horizon) #same for upper and lower intervals
fLowerHolder = matrix(nrow = n_windows, ncol = horizon)
phis = m1$phi
thetas = m1$theta
s = 0
d = 0

for( i in 1:n_windows) #how many "windows" can fit into the total length of time, rounding down
{
  t_start = 1+(step_size*(i-1)) #starting point for each window's training data
  t_end = trainingSize + (step_size*(i-1)) #endpoint
  forecasts = fore.aruma.wge(tdata[t_start:t_end],phi = phis, theta = thetas, s = s, d = d, n.ahead = horizon)
  
  ASE = mean((tdata[(t_end + 1):(t_end+horizon)] - forecasts$f)^2)
  ASEHolder[i] = ASE
  fcastHolder[i, ] <- forecasts$f
  fUpperHolder[i, ] <- forecasts$ul
  fLowerHolder[i, ] <- forecasts$ll
  #print(paste(t_start, t_end))
}

ASEHolder
WindowedASE = mean(ASEHolder)
hist(ASEHolder, main = paste("Mean ASE across [", n_windows, '] moving windows: ', WindowedASE))

summary(ASEHolder)
print(paste("Mean ASE across ", n_windows, ' moving windows: ', WindowedASE))
```

```{r}
#function to space out forecasts with NAs for plotting
fspacer <- function(fcasts){
  cast_frame = 0
  cast_frame = matrix(nrow = n_windows, ncol = length(tdata)) #empty matrix to hold spaced out forecasts
  for(i in 1:n_windows)
    {
    #print(i)
    before_block <- rep(NA, (trainingSize + (step_size*(i-1))))
    after_block <- c(rep(NA, (length(tdata) - length(c(rep(NA, (trainingSize + step_size*(i-1))))) - horizon)))
    #print(paste(length(before_block), length(after_block)))
    #print(length(c(before_block, fcasts[i, ], after_block)))
    cast_frame[i, ] <- c(before_block, fcasts[i, ], after_block)
    #print(cast_frame[i, ])
    }
  return(cast_frame)
}
f_spaced <- fspacer(fcastHolder)
```

```{r}
#create categorical dataframe to store windowed predictions
fframe <- list()
for(i in 1:n_windows){
  #name = paste0('df', i)
  lname = paste0('w', i)
  #assign(name, data.frame(date = ttime, pred = f_spaced[i, ]))
  tmp <- data.frame(date = ttime, pred = f_spaced[i, ])
  fframe[[lname]] <- tmp
  }
f_df <- cbind(cat= rep(names(fframe), sapply(fframe, NROW)), do.call(rbind, fframe)) #arrange data with category by prediction window
o_df <- data.frame(cat = rep('original', length(tdata)), date = ttime, pred = tdata) #add actual data as rbind
f_df <- rbind(f_df, o_df)

ggplot(f_df, aes(date,pred, color= cat)) + geom_line() +
  ggtitle("Total Influenza Test Cases: India 2017 - 2019", subtitle = paste("Mean ASE across [", n_windows, '] moving windows: ', WindowedASE))

```






















